---
title: "Assignment 1"
author: "Warren"
date: "13/04/2020"
output: html_document
---

## 1. Reading in the data, forked from course website
```{r}
unzip("C:/Users/Herman/Documents/MDRP/RepData_PeerAssessment1/activity.zip")

MOVE<- read.csv("C:/Users/Herman/Documents/MDRP/RepData_PeerAssessment1/Activity.csv")
```

## 2. What is the total number of steps per day?

### Looking at data set
```{r}
head(MOVE)
names(MOVE)
dim(MOVE)
summary(MOVE)
```

### Calculating Total Number of Steps per Day
```{r}
TotalSPD<- tapply(MOVE$steps, MOVE$date, sum)
TotalSPD
```

### Drawing Histogram
```{r, histogram, fig.height=4}
library(ggplot2)
qplot(TotalSPD, binwidth=500, main="Total number of steps per day")
```

### Calculating mean and median of the total no. of steps each day
```{r}
mean(TotalSPD, na.rm=TRUE)
median(TotalSPD, na.rm=TRUE)
```

## 3. What is the average daily activity pattern? 

### Calculate Average no. of steps for each interval
```{r}
mean_interval<-tapply(MOVE$steps, MOVE$interval, mean, na.rm=TRUE)
```

### Draw Time Series
```{r, line graph, fig.height=4}
average<-as.data.frame(mean_interval)
intervals <-as.data.frame(unique(MOVE$interval))
df<-data.frame(average,intervals)
colnames(df)<- c("average", "intervals")
qplot(df$intervals, df$average, 
      geom="line", main="TIME SERIES", xlab="5-minute intervals", 
      ylab="Average steps")
```

### Finding which day contains the maximum number of steps
```{r}
df[which.max(df$average),]
```

## 4.Imputing missing values

### calculate no. of missing values
```{r}
sum(is.na(MOVE$steps))
```
      
### Let missing value = mean for that interval
#### Subset data where there are no NAs
```{r}
MOVE_no_NA<- MOVE[which(!is.na(MOVE$steps)),]
```

####Convert average per interval into intefers
```{r}
df$average<- as.integer(df$average)
```

#### Subset data where steps have NAs
```{r}
MOVE_NA<-MOVE[which(is.na(MOVE$steps)),]
```
            
#### Fill NAs with average steps based on interval
```{r}
MOVE_NA$steps<- ifelse(MOVE_NA$interval==df$intervals, df$average)
```
            
#### Row bind datasets such that NAs are replaced with means
```{r}
MOVE_impute<- rbind(MOVE_no_NA, MOVE_NA)
```

### Drawing historgram of total number of steps taken daily
```{r}
      TotalSPD_impute<- tapply(MOVE_impute$steps, MOVE_impute$date, sum)
      qplot(TotalSPD_impute, binwidth=500, main="Total number of steps per day")
```

### Calculating mean and median      
```{r}
mean(TotalSPD_impute, na.rm=TRUE)
median(TotalSPD_impute, na.rm=TRUE)
```
      
## 5. Differences in activity patterns on weekdays vs weekends

### Changing dates to days of the week
```{r}
MOVE_impute$day<-weekdays(as.Date(MOVE_impute$date))
```

### Load dplyr and group by day
```{r}
library(dplyr)
group_by(MOVE_impute, day)
```
      
### Changing day of week to weekday/weekend
```{r}
MOVE_impute$day<- ifelse(MOVE_impute$day==
            c("Saturday", "Sunday"), yes= "Weekend", no="Weekday")
```

### Making panel plot
```{r, histogram2, fig.height=4}
library(dplyr)
meansteps <- MOVE_impute %>% group_by(interval,day) %>% summarise(average = mean(steps))

qplot(interval,average,data=meansteps,
            geom="line",facets=day~.,xlab="5-minute interval",
            ylab="average number of steps",
            main="Average steps pattern between Weekday and Weekend")
```
      
            
      
